\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-06-07 nampi.tex
shortened

\section{Overview of \system}

We take inspiration from several ways that skilled human
programmers have learned to code: Skilled coders build libraries of
reusable subroutines that are shared across related programming tasks,
and can be composed to generate increasingly complex and powerful
subroutines.  In text editing, a good library should support routines
for splitting on characters, but also specialize these routines to
split on frequent delimiters such as spaces or commas. Skilled coders
also learn to recognize what kinds of programming idioms and library
routines would be useful for solving the task at hand, even if they
cannot instantly work out the details.  In text editing, one might
learn that if outputs are consistently shorter than inputs, removing
characters is likely to be part of the solution.%% ; if every output
%% contains a constant substring (e.g., ``Dr.''), inserting or appending
%% that constant string is likely to be a subroutine.

@@

%% Our \system
%% algorithm incorporates these insights by iterating
Concretely, our algorithm iterates through three steps. 
The \textbf{Wake} step takes a given set of \textbf{tasks}, typically
several hundred, and searches for compact programs that solve these tasks
guided by the current DSL and neural network.
The \textbf{Sleep-G} step grows the DSL (or \textbf{G}enerative model) 
which allows the agent to more compactly
write programs in the domain; it modifies the structure of the DSL by
discovering regularities across programs found during waking, compressing
them to distill out common code fragments across successful programs.
The \textbf{Sleep-R} step improves the search procedure by training a neural network (the \textbf{R}ecognition model) to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference}. %%  (CITE ,
%% and others too on amortized inference.)
%Crucially, these steps bootstrap off each other~(Figure~\ref{feeding}).
%% \begin{figure}[h]\centering
%%   \begin{tikzpicture}
%%     \begin{scope}[shift = {(1,-1)}]
%%     \node[align = center](synthesis) at (6,4) {Search for \\programs};
%%     \node[align = center](DSL) at (3,1) {DSL};
%%     \node[align = center](recognitionModel) at (9,1) {Neural\\recognition \\model};

@@

Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}),
neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
which also learns a library of subroutines.
We find EC motivating, and go beyond it and other prior work through the following contributions: %% \begin{itemize}
%%   \setlength\itemsep{0}
%%   \item We show how to learn-to-learn programs in an expressive Lisp-like programming language;
%%   \item We give an algorithm for learning DSLs, built on a formalism known
%%     as Fragment Grammars~\cite{tim}; and 
%%   \item We give a hierarchical Bayesian framing of the problem that allows a
%%     joint inference of the DSL and recognition model.
%% \end{itemize}
\\\textbf{Contributions.} (1) We show how to learn-to-learn programs in an expressive Lisp-like programming language, including conditionals, variables, and higher-order recursive functions;
(2) We give an algorithm for learning DSLs,
built on a formalism known as Fragment Grammars~\cite{tim};
and (3) We give a hierarchical Bayesian framing enabling joint inference of the DSL and recognition model.

>>>>

%\section{Overview of \system}

We take inspiration from several ways that skilled human
programmers have learned to code: Skilled coders build libraries of
reusable subroutines that are shared across related programming tasks,
and can be composed to generate increasingly complex
subroutines.  In text editing, a good library should support routines
for splitting on characters, but also specialize these routines to
split on frequent delimiters such as spaces or commas. Skilled coders
also learn to recognize what kinds of programming idioms and library
routines would be useful for solving the task at hand, even if they
cannot instantly work out the details.  In text editing, one might
learn that if outputs are consistently shorter than inputs, removing
characters is likely part of the solution.%% ; if every output
%% contains a constant substring (e.g., ``Dr.''), inserting or appending
%% that constant string is likely to be a subroutine.

@@

%% Our \system
%% algorithm incorporates these insights by iterating
Concretely, our algorithm iterates through three steps. 
The \textbf{Wake} step takes a given set of \textbf{tasks}, typically
several hundred, and searches for compact programs that solve these tasks,
guided by the current DSL and neural network.
The \textbf{Sleep-G} step grows the DSL (or \textbf{G}enerative model) 
which allows the agent to more compactly
write programs in the domain; it modifies the structure of the DSL by
discovering regularities across programs found during waking, compressing
them to distill out common code fragments across successful programs.
The \textbf{Sleep-R} step improves the search procedure by training a neural network (the \textbf{R}ecognition model) to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference}. %%  (CITE ,
%% and others too on amortized inference.)
%Crucially, these steps bootstrap off each other~(Figure~\ref{feeding}).
%% \begin{figure}[h]\centering
%%   \begin{tikzpicture}
%%     \begin{scope}[shift = {(1,-1)}]
%%     \node[align = center](synthesis) at (6,4) {Search for \\programs};
%%     \node[align = center](DSL) at (3,1) {DSL};
%%     \node[align = center](recognitionModel) at (9,1) {Neural\\recognition \\model};

@@

\textbf{Related work.} Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}),
neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
which also learns a library of subroutines.
We find EC motivating, and go beyond it and other prior work through the following contributions: %% \begin{itemize}
%%   \setlength\itemsep{0}
%%   \item We show how to learn-to-learn programs in an expressive Lisp-like programming language;
%%   \item We give an algorithm for learning DSLs, built on a formalism known
%%     as Fragment Grammars~\cite{tim}; and 
%%   \item We give a hierarchical Bayesian framing of the problem that allows a
%%     joint inference of the DSL and recognition model.
%% \end{itemize}
\\\textbf{Contributions.} (1) We show how to learn-to-learn programs in an expressive Lisp-like programming language, including conditionals, variables, and higher-order recursive functions;
(2) We give an algorithm for learning DSLs,
built on a formalism known as Fragment Grammars~\cite{tim};
and (3) We give a hierarchical Bayesian framing enabling joint inference of the DSL and recognition model.
